---
title: "Analysis"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r loadPackages, message=FALSE, warning=FALSE}
library(dplyr)
library(MachineShop)
library(recipes)
library(kableExtra)
library(arsenal)
```

```{r controlSetup, results='hide', echo=FALSE}
library(doSNOW)
registerDoSNOW(makeCluster(8))
## Resample control object
ctrl <- CVControl(seed = 123)
```


```{r preProcess, echo = FALSE}
dat <- readRDS("./data/hcc_data.RDS")

rec_base <- recipe(
  status ~ .,
  data = dat
) %>%
  role_case(stratum = status) %>%
  check_missing(status) %>%
  step_num2factor(status, transform = function(x) x + 1,
                  levels = c("died", "survive")) %>%
  step_num2factor(gender, transform = function(x) x + 1,
                  levels = c("female", "male")) %>%
  step_num2factor(symptom, transform = function(x) x + 1,
                  levels = c("no", 'yes')) %>%
  step_num2factor(alc, transform = function(x) x + 1,
                  levels = c("no", 'yes')) %>%
  step_num2factor(hepBsurfAnti,transform = function(x) x + 1,
                  levels = c("no", 'yes')) %>%
  step_num2factor(hepBeAnti,transform = function(x) x + 1,
                  levels = c("no", 'yes')) %>%
  step_num2factor(hepBcorAnti,transform = function(x) x + 1,
                  levels = c("no", 'yes')) %>%
  step_num2factor(hepCvirAnti,transform = function(x) x + 1,
                  levels = c("no", 'yes')) %>%
  step_num2factor(cirr,transform = function(x) x + 1,
                  levels = c("no", 'yes')) %>%
  step_num2factor(endemicCountries,transform = function(x) x + 1,
                  levels = c("no", 'yes')) %>%
  step_num2factor(smoke,transform = function(x) x + 1,
                  levels = c("no", 'yes')) %>%
  step_num2factor(diabetes,transform = function(x) x + 1,
                  levels = c("no", 'yes')) %>%
  step_num2factor(obese,transform = function(x) x + 1,
                  levels = c("no", 'yes')) %>%
  step_num2factor(hemochro,transform = function(x) x + 1,
                  levels = c("no", 'yes')) %>%
  step_num2factor(artHyper,transform = function(x) x + 1,
                  levels = c("no", 'yes')) %>%
  step_num2factor(chronRenal,transform = function(x) x + 1,
                  levels = c("no", 'yes')) %>%
  step_num2factor(hiv,transform = function(x) x + 1,
                  levels = c("no", 'yes')) %>%
  step_num2factor(Nasteato,transform = function(x) x + 1,
                  levels = c("no", 'yes')) %>%
  step_num2factor(esophVarices,transform = function(x) x + 1,
                  levels = c("no", 'yes')) %>%
  step_num2factor(spleno,transform = function(x) x + 1,
                  levels = c("no", 'yes')) %>%
  step_num2factor(portalHyper,transform = function(x) x + 1,
                  levels = c("no", 'yes')) %>%
  step_num2factor(portalVeinThromb,transform = function(x) x + 1,
                  levels = c("no", 'yes')) %>%
  step_num2factor(LiverMeta,transform = function(x) x + 1,
                  levels = c("no", 'yes')) %>%
  step_num2factor(RadioHallmark,transform = function(x) x + 1,
                  levels = c("no", 'yes')) %>%
  step_num2factor(numNodules,transform = function(x) x + 1,
                  levels = as.character(0:5))

dat <- bake(prep(rec_base), new_data = dat)
```

```{r summary, results='asis', echo = FALSE}
tbl <- tableby(
  ~ ., data = dat,
  control = tableby.control(
    cat.stats = c("countpct", "Nmiss"),
    numeric.stats = c("medianrange", "Nmiss"),
    digits = 2
  )
)

summary(tbl, title = 'HCC Data Summary') 
# %>% as.data.frame()
  # kable_styling(full_width=FALSE)
```


```{r fitting}
# dir.create("./data/ModelCache/")
fnames <- paste0("./data/ModelCache/",
                 c("lasso_fit.RDS", "lasso_res.RDS"))
### Imputation step
rec <- rec_base %>%
  step_impute_mean(all_numeric_predictors()) %>%
  step_impute_mode(all_nominal_predictors())

model <- TunedModel(GLMNetModel(alpha = 1),
                    grid = c(lambda = 10))

if(file.exists(fnames[1])){
  model_fit <- readRDS(fnames[1])
}else{
  model_fit <- fit(rec, model, control = ctrl)
  saveRDS(model_fit, file = fnames[1])
}

if(file.exists(fnames[2])){
  model_res <- readRDS(fnames[2])
}else{
  model_res <- resample(rec, model, control = ctrl)
  saveRDS(model_res, file = fnames[2])
}

summary(model_res)
plot(model_res)
cf <- coef(unMLModelFit(model_fit))

rownames(cf)[cf[,1] !=0]

### nonzero variables from lasso fit

nz_cf <- 
c("symptom",
  "hepBsurfAnti",
  "hepCvirAnti",
  "endemicCountries",
  "smoke",
  "diabetes",
  "artHyper",
  "Nasteato",
  "portalHyper",
  "ascites",
  "international",
  "alphaFreto",
  "hemaglobin",
  "leukocyte",
  "portalVeinThromb",
  "LiverMeta",
  "Age",
  "cigPerYr",
  "performance",
  "albumin",
  "alanineTransaminase",
  "aspartateTransaminase",
  "alkalinePhosphatase",
  "totalProteins",
  "creatine",
  "numNodules",
  "majorDimNode",
  "directBilirubin",
  "iron",
  "ferritin")


rec2 <- recipe(
  as.formula(paste0("status ~ ", paste0(nz_cf, collapse = '+'))),
  data = dat) %>%
  role_case(stratum = status) %>%
  check_missing(status) %>%
  step_impute_mean(all_numeric_predictors()) %>%
  step_impute_mode(all_nominal_predictors())

glm_fit <- fit(rec2, model =GLMModel)
varimp(glm_fit, method = 'model', scale = FALSE) %>% plot()
```


```{r}
set.seed(123123)
test1 <- varimp(model_fit, scale = FALSE, stats = mean)
test1 <- test1[order(rownames(test1)),]
plot(test1)
set.seed(123123)
test2 <- varimp(model_fit, scale = FALSE, stats = function(x){mean(x > 0)})
test2 <- test2[order(rownames(test2)),]
plot(test2)
```

Data Transformations
-different of vectors
-pca transformation

Imputation Steps
-mean and mode imputation
-step impute bag() FN Imputation via bagged trees
-step impute knn() FN Imputation via K-nearest neighbors
-step impute lower()
-multiple impuation (MICE) not a step

Filtering Variables
-step corr() N High correlation filter
-step lincomb()
-step nzv() FN Near-zero variance filter
-Distribution step zv()

Model Algorithms

-nnet josh
-xgboost/gbm/adaboost yujing
-randomforest sangil
-stacked regression
-logistic
-lasso/ridge
