---
title: "Analysis"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r checkTemplate, echo=FALSE}
exists("fname_base")
exists("mspec")
exists("rec")
save(fname_base, mspec, rec, file = paste0(fname_base, "_template_params.rda"))
# load(file = "knn_NNet_template_params.rda")
```

```{r loadPackages, message=FALSE, warning=FALSE}
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(MachineShop))
suppressPackageStartupMessages(library(recipes))
suppressPackageStartupMessages(library(kableExtra))
suppressPackageStartupMessages(library(arsenal))
```

```{r controlSetup, results='hide', echo=FALSE}
suppressPackageStartupMessages(library(doSNOW))
registerDoSNOW(makeCluster(parallel::detectCores()))
## Resample control object
# ctrl <- CVControl(seed = 123)
```


```{r fitting}
# dir.create("./data/ModelCache/")
fnames <- paste0("../data/ModelCache/",
                 paste0(fname_base, c("_fit.RDS", "_res.RDS"))
                )

if(file.exists(fnames[1])){
  model_fit <- readRDS(fnames[1])
}else{
  model_fit <- fit(mspec, control = ctrl)
  saveRDS(model_fit, file = fnames[1])
}

if(file.exists(fnames[2])){
  model_res <- readRDS(fnames[2])
}else{
  model_res <- resample(mspec, control = ctrl)
  saveRDS(model_res, file = fnames[2])
}

# model_fit <- as.MLModel(model_fit)
varimp(model_fit) %>% plot()
summary(model_res)
plot(model_res)
confusion(model_res)
### nonzero variables from lasso fit

```

Data Transformations
-different of vectors
-pca transformation

Imputation Steps
-mean and mode imputation
-step impute bag() FN Imputation via bagged trees
-step impute knn() FN Imputation via K-nearest neighbors
-step impute lower()
-multiple impuation (MICE) not a step

Filtering Variables
-step corr() N High correlation filter
-step lincomb()
-step nzv() FN Near-zero variance filter
-Distribution step zv()

Model Algorithms

-nnet josh
-xgboost/gbm/adaboost yujing
-randomforest sangil
-stacked regression
-logistic
-lasso/ridge
